Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
Adds ww data
Check out new results
Look, @mariosky, they are so much better than what we had before. Great job!

It is interesting also to check the distribution of the experiment
duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
a Zipf's law, with similar distribution along all three runs. The 4/24
run is the most complete and shows a S-shape, which implies an
accumulation of experiments taking similar time and around 100
seconds. The most interesting part is the {\em tail}, which shows how
many experiments took a desirable amount of time, and of the order of
10 seconds. As it can be seen, it drops sharply implying there are
just a few of them, and with diminishing probability as time
decreases. This exponential distribution also appears in the
distribution of HTTP {\tt PUT}s, equivalent to the number of
generations divided by 100, contributed by every user, which is shown
in Figure \ref{fig:put:os}. These results show a Zipf-like behavior,
so we have fitted it to the Generalized Extreme Value distribution,
with the resulting parameters shown in Table \ref{tab:put:os}.
%
\begin{table*}
\caption{Summary of fit to Generalized Extreme Value distribution of
  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
\begin{center}
\begin{tabular}{l|ccc}
\hline
Date  & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
\hline 
4/4 &  8.541 $\pm$ 1.0926  &    12.442 $\pm$ 1.7302 &  1.388 $\pm$
0.1377 \\  
4/24 & 6.148 $\pm$ 0.3782 & 7.354 $\pm$ 0.5105 & 1.090 $\pm$  0.0697  \\    
7/31 & 11.645 $\pm$ 1.475 & 16.365 $\pm$ 2.201 &  1.265 $\pm$ 0.132   \\   
\hline 
\end{tabular}
\end{center}
\end{table*}
%
This distribution was originally proposed to fit extreme values
\cite{resnick2013extreme} and contains, as a special case, the inverse
Weibull distribution which was fitted to volunteer computing
frameworks such as SETI@home \cite{javadi2009mining}. This
distribution is governed by three parameters, the usual location $\mu$
which is related to where it has its {\em center} and  scale $\sigma$,
related to the size, but also a third shape $\xi$ parameter that is
related to its skewness, that is, how skewed it is around the central
location. Positive parameters indicate that the distribution {\em
  leans} towards the origin, negative towards the other extreme
value. In this case, Table \cite{tab:puts:os} shows $\xi$ values
greater than one and between one and 1.4, which indicates that the
three experiment share this origin-leaning pattern, with many users
donating a few cycles and just a few donating extreme values. Random
distributions with these parameters have been plotted in red in
\ref{fig:put:os}, indicating that the fit is good enough. The
predictive value of these fits, however, is limited, over all taking
into account the low correlation between successive events shown
above.

@@

The sequence of events are:
\begin{enumerate}
\item First the NodIO process is started in nodejs. As part of the 
initialization procedure the number of experiment is initialised and a log 
file is created. The shared pool implemented as an array is reset.
\item A volunteer follows the link of the experiment triggering a GET HTTP 
request and the page is loaded.
\item The main script renders the page and creates the worker instance. In 
order to create a dedicated worker instance another GET request is sent to the 
server in order to recieve the worker implementation script. 
Each worker instance executes one or more evolutionary algorithms 
{\em islands}. Each starts with random population and it is assigned a 
universally unique  identifier (UUID). The UUID is included in each of the HTTP
requests sent to the server.  
The initial paramaters for each island
can be distinct. Once the worker is created and the parameters have been set,
a message is sent to the worker to start the algorithm.
\item The worker first creates the population following with the previously 
specified parameters. Once the population has been initialised, the EA 
generation loop is started. After n generations the state of the population 
is evaluated. First the chromosome with the best fitness is sent to the shared
population with a PUT HTTP request. If the best fitness is not a global best,
the EA loop must continue, but first a chromosome is randomly selected from 
the shared pupulation. Also the current generation number and the best 
chromosome is sent to the client main thread in order to update the DOM to 
display the current state of the island (described next). These messages are
all asynchronous so even before a message is recieved by target the EA loop of
the worker could continue. If the best fitness is a global best the best 
chromosme is sent to the server and the loop ends. A message is sent to the 
main thread indicating that the best individual has been found.
\item The main thread recieves messages from workers as events handled by 
a callback function. When an iteration message is recived ,the page is 
updated by displaying a dinamic plot of the geneartion number and best 
fitness, and a representation of the chromosme is also generated.   
\item When a global best is recieved from an island the current experiment 
ends, the experiment number is incremented and the population array is reset.  
\item When the main thread recieves a message the a worker has ended the EA
loop, that worker is reinitialized. The worker process is not ended, to
avoid the cost, only the parameters and population is reset and a new UIID 
assigned.
\end{enumerate}

\begin{figure}[!t]
\centering
\includegraphics[width=4in]{img/Algorithm.pdf}
\caption{Description of the W$^2$ version of {\sf NodIO}. In this
  version, clients use Web Workers to run the evolutionary algorithm,
  with two of them per browser.}  
\label{fig:system:w2}
\end{figure}

%---------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}

>>>>

It is interesting also to check the distribution of the experiment
duration, shown in Figure \ref{fig:zipf:os} and which roughly follows
a Zipf's law, with similar distribution along all three runs. The 4/24
run is the most complete and shows a S-shape, which implies an
accumulation of experiments taking similar time and around 100
seconds. The most interesting part is the {\em tail}, which shows how
many experiments took a desirable amount of time, and of the order of
10 seconds. As it can be seen, it drops sharply implying there are
just a few of them, and with diminishing probability as time
decreases. This exponential distribution also appears in the
distribution of HTTP {\tt PUT}s, equivalent to the number of
generations divided by 100, contributed by every user, which is shown
in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
so we have fitted it to the Generalized Extreme Value distribution,
with the resulting parameters shown in Table \ref{tab:puts:os}.
%
\begin{table*}
\caption{Summary of fit to Generalized Extreme Value distribution of
  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
\begin{center}
\begin{tabular}{l|ccc}
\hline
Date  & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
\hline 
4/4 &  8.541 $\pm$ 1.0926  &    12.442 $\pm$ 1.7302 &  1.388 $\pm$
0.1377 \\  
4/24 & 6.148 $\pm$ 0.3782 & 7.354 $\pm$ 0.5105 & 1.090 $\pm$  0.0697  \\    
7/31 & 11.645 $\pm$ 1.475 & 16.365 $\pm$ 2.201 &  1.265 $\pm$ 0.132   \\   
\hline 
\end{tabular}
\end{center}
\end{table*}
%
This distribution was originally proposed to fit extreme values
\cite{resnick2013extreme} and contains, as a special case, the inverse
Weibull distribution which was fitted to volunteer computing
frameworks such as SETI@home \cite{javadi2009mining}. This
distribution is governed by three parameters, the usual location $\mu$
which is related to where it has its {\em center} and  scale $\sigma$,
related to the size, but also a third shape $\xi$ parameter that is
related to its skewness, that is, how skewed it is around the central
location. Positive parameters indicate that the distribution {\em
  leans} towards the origin, negative towards the other extreme
value. In this case, Table \ref{tab:puts:os} shows $\xi$ values
greater than one and between one and 1.4, which indicates that the
three experiment share this origin-leaning pattern, with many users
donating a few cycles and just a few donating extreme values. Random
distributions with these parameters have been plotted in red in
\ref{fig:puts:os}, indicating that the fit is good enough. The
predictive value of these fits, however, is limited, over all taking
into account the low correlation between successive events shown
above.

@@

The sequence of events are:
\begin{enumerate}
\item First the NodIO process is started in nodejs. As part of the 
initialization procedure the number of experiment is initialised and a log 
file is created. The shared pool implemented as an array is reset.
\item A volunteer follows the link of the experiment triggering a GET HTTP 
request and the page is loaded.
\item The main script renders the page and creates the worker instance. In 
order to create a dedicated worker instance another GET request is sent to the 
server in order to recieve the worker implementation script. 
Each worker instance executes one or more evolutionary algorithms 
{\em islands}. Each starts with random population and it is assigned a 
universally unique  identifier (UUID). The UUID is included in each of the HTTP
requests sent to the server.  
The initial paramaters for each island
can be distinct. Once the worker is created and the parameters have been set,
a message is sent to the worker to start the algorithm.
\item The worker first creates the population following with the previously 
specified parameters. Once the population has been initialised, the EA 
generation loop is started. After n generations the state of the population 
is evaluated. First the chromosome with the best fitness is sent to the shared
population with a PUT HTTP request. If the best fitness is not a global best,
the EA loop must continue, but first a chromosome is randomly selected from 
the shared pupulation. Also the current generation number and the best 
chromosome is sent to the client main thread in order to update the DOM to 
display the current state of the island (described next). These messages are
all asynchronous so even before a message is recieved by target the EA loop of
the worker could continue. If the best fitness is a global best the best 
chromosme is sent to the server and the loop ends. A message is sent to the 
main thread indicating that the best individual has been found.
\item The main thread recieves messages from workers as events handled by 
a callback function. When an iteration message is recived ,the page is 
updated by displaying a dinamic plot of the geneartion number and best 
fitness, and a representation of the chromosme is also generated.   
\item When a global best is recieved from an island the current experiment 
ends, the experiment number is incremented and the population array is reset.  
\item When the main thread recieves a message the a worker has ended the EA
loop, that worker is reinitialized. The worker process is not ended, to
avoid the cost, only the parameters and population is reset and a new UIID 
assigned.
\end{enumerate}
%
\begin{figure}[!t]
\centering
\includegraphics[width=4in]{img/Algorithm.pdf}
\caption{Description of the W$^2$ version of {\sf NodIO}. In this
  version, clients use Web Workers to run the evolutionary algorithm,
  with two of them per browser.}  
\label{fig:system:w2}
\end{figure}
%

\begin{table*}
\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline 
{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
\hline 
\end{tabular}
\end{center}
\end{table*}
%
%---------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
